---
title: "Assignment 1"
subtitle: "Identify a dataset"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Find or assemble a dataset with

* At least three continuous variables
* At least one categorical (or ordinal variables)
* Between 100 and 1 million observations

Write out your answers to the following questions:

1. What is the question you will seek to answer through your analysis 
of the data?
2. Which variable is your outcome/dependent/response variable? (Note: this *must* be a continuous variable) What does this variable represent?
3. Which variables are continuous variables? What does each variable represent?
4. Which variables are categorical or ordinal variables (specify which)? What does each variable represent? 
5. What is your sample size?
6. What is the sampling method?
7. What is your population of interest?
8. What is your unit of analysis?
9. What is your data source?

## Deliverable

Submit a csv file with your dataset and a document with your answers to the above questions.

## Grading criteria

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)

rubric <- tibble(Item = c("Identification of a research question",
                          "~",
                          "Identification of an outcome variable",
                          "~",
                          "Identification of continuous variables",
                          "~",
                          "Identification of a categorical/ordinal variable",
                          "~",
                          "Statement of sample size",
                          "~",
                          "Description of sampling method",
                          "~",
                          "Identification of population of interest",
                          "~",
                          "Identification of unit of analysis",
                          "~",
                          "Identification of data source",
                          "~",
                          "Submission of dataset",
                          "~",
                          "Total"),
                 `Half credit` = 
                   c("Inconsistent with data",
                     "~",
                     "Misunderstanding of outcome variable meaning, inconsistent with research question, and/or not continuous",
                     "~",
                     "Not continuous variables, inconsistent with research question, inconsistent with data, and/or unclear description",
                     "~",
                     "Not categorical or ordinal variables, inconsistent with data, and/or unclear description",
                     "~",
                     "Inconsistent with data.",
                     "~",
                     "Unclear relationship between sample and population of interest",
                     "~",
                     "Unclear or unreasonable",
                     "~",
                     "Inconsistent with data",
                     "~",
                     "Unclear", 
                     "~",
                     "All or nothing, no half credit",
                     "~",""),
                 `Full credit` = c("Consistent with data",
                                   "~",
                                   "Clear understanding of outcome variable meaning,  consistent with research question, and variable is continuous",
                                   "~",
                                   "Continuous variables, consistent with research question and data, and clear description",
                                   "~",
                                   "Correctly identified categorical or ordinal variables, consistent with data, clear description",
                                   "~",
                                   "Consistent with data",
                                   "~",
                                   "Clear relationship between sample and population of interest",
                                   "~",
                                   "Clear and reasonable",
                                   "~",
                                   "Consistent with data",
                                   "~",
                                   "Clear",
                                   "~",
                                   "Submitted a dataset",
                                   "~",
                                   ""),
                 `Possible points` = c(2, "",
                                       2,"",
                                       2,"",
                                       2,"",
                                       2,"",
                                       2,"",
                                       2,"", 
                                       2,"",
                                       7, "",
                                       0,"",
                                       25))

kable(rubric, format = "html")
```
        