---
title: "Assignment 1"
subtitle: "Identify a dataset"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Find or assemble a dataset with

* At least three continuous variables
* At least one categorical (or ordinal variables)
* Between 100 and 1 million observations

Answer the following questions:
    
1. Which variable is your outcome/dependent/response variable? (Note: this *must* be a continuous variable) What does this variable represent?
2. Which variables are continuous variables? What does each variable represent?
3. Which variables are categorical or ordinal variables (specify which)? What does each variable represent? 
4. What is your sample size?
5. What is the sampling method?
6. What is your population of interest?
7. What is your unit of analysis?
8. What is your data source?
9. (optional) Name any classmates you know of who are working on something similar.

Write out a research question in the following form: 

*"What is the effect of a* **[unit of analysis]***'s* **[variable(s) of interest]** *on its* 
**[outcome variable]***, after accounting for the effects of* **[list of control variables]***?"*

## Deliverable

Submit a document with your research question (in the specified format) and your
answers to the above questions. Submit your dataset as a csv file.

## Examples

Coming soon.

## Grading criteria

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)

rubric <- tibble(Item = c("Identification of a research question",
                          "~",
                          "Identification of an outcome variable",
                          "~",
                          "Identification of continuous variables",
                          "~",
                          "Identification of a categorical/ordinal variable",
                          "~",
                          "Statement of sample size",
                          "~",
                          "Description of sampling method",
                          "~",
                          "Identification of population of interest",
                          "~",
                          "Identification of unit of analysis",
                          "~",
                          "Identification of data source(s)",
                          "~",
                          "Total"),
                 `Half credit` = 
                   c("Does not follow the recommended format",
                     "~",
                     "Misunderstanding of outcome variable meaning, inconsistent with research question, and/or not continuous",
                     "~",
                     "Not continuous variables, inconsistent with research question, and/or unclear description",
                     "~",
                     "Not categorical or ordinal variables and/or unclear description",
                     "~",
                     "Not between 100 and 1,000,000",
                     "~",
                     "Unclear relationship between sample and population of interest",
                     "~",
                     "Unclear or unreasonable",
                     "~",
                     "Inconsistent with research question",
                     "~",
                     "Unclear", 

                     "~",""),
                 `Full credit` = c("Follows recommended formal",
                                   "~",
                                   "Clear understanding of outcome variable meaning,  consistent with research question, and variable is continuous",
                                   "~",
                                   "Continuous variables, consistent with research question, and clear description",
                                   "~",
                                   "Correctly identified categorical or ordinal variables, clear description",
                                   "~",
                                   "Between 100 and 1,000,000",
                                   "~",
                                   "Clear relationship between sample and population of interest",
                                   "~",
                                   "Clear and reasonable",
                                   "~",
                                   "Consistent with research question",
                                   "~",
                                   "Clear",
                                   "~",
                                   ""),
                 `Possible points` = c(9, "",
                                       2,"",
                                       2,"",
                                       2,"",
                                       2,"",
                                       2,"",
                                       2,"", 
                                       2,"",
                                       2, "",
                                       25))

kable(rubric, format = "html")
```
        