---
title: "Assignment 2"
subtitle: "Visualizing data"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Calculate the following values for each continuous variable in your 
data set:

* Average
* Median
* Standard deviation
* Minimum
* Maximum
* 25th percentile
* 75th percentile

Create a figure that shows the distribution of each continuous variable.

Create a figure that shows the share of each categorical variable in each category.

## Deliverable

Submit a memo (as a PDF or html document) presenting and explaining your figures and 
the results of your calculations.

## Example

These examples would receive full credit.

[Commuting example](https://c-voulgaris.github.io/SES-5215-F22/Examples/a2-commuting.pdf){target="_blank"}

[Crashes example](https://c-voulgaris.github.io/SES-5215-F22/Examples/a2-crashes.pdf){target="_blank"}

## Grading criteria

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)

rubric <- tibble(Item = c("Calculated values",
                          "~",
                          "Distribution figures",
                          "~",
                          "Proportion figures",
                          "~",
                          "Average within category figures",
                          "~",
                          "Presentation of calculated values",
                          "~",
                          "Explanation of calculated values",
                          "~",
                          "Explanation of figures",
                          "~",
                          "Total"),
                 `Half credit` = 
                   c("Unreasonable values",
                     "~",
                     "Distribution not clear",
                     "~",
                     "Proportions not clear",
                     "~",
                     "Averages not illustrated",
                     "~",
                     "Values not clearly shown",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unclear explanation",
                     "~",""),
                 `Full credit` = c("Reasonable values",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear presentation",
                                   "~",
                                   "Clear explanation",
                                   "~",
                                   "Clear explanation",
                                   "~",
                                   ""),
                 `Possible points` = c(5, "~",
                                       3,"~",
                                       3,"~",
                                       3,"~",
                                       3,"~",
                                       4,"~", 
                                       4,"~",
                                       25))

kable(rubric, format = "html")
```
        