---
title: "Assignment 2"
subtitle: "Visualizing data"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Calculate the following values for each continuous variable in your 
data set:

* Average
* Median
* Standard deviation
* Minimum
* Maximum
* 25th percentile
* 75th percentile

Create a figure that shows the distribution of each continuous variable.

Create a figure that shows the share of each categorical variable in each category.

Create a figure that shows the relationship of each independent continuous
variable with your dependent (outcome) variable.

Create a figure that shows the average value of your outcome variable within
each category of each categorical variable.

## Deliverable

Submit a memo (as a PDF or html document) presenting and explaining your figures and 
the results of your calculations.

## Grading criteria

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)

rubric <- tibble(Item = c("Calculated values",
                          "~",
                          "Distribution figures",
                          "~",
                          "Proportion figures",
                          "~",
                          "Continuous relationship figures",
                          "~",
                          "Average within category figures",
                          "~",
                          "Presentation of calculated values",
                          "~",
                          "Explanation of calculated values",
                          "~",
                          "Explanation of figures",
                          "~",
                          "Total"),
                 `Half credit` = 
                   c("Unreasonable values",
                     "~",
                     "Distribution not clear",
                     "~",
                     "Proportions not clear",
                     "~",
                     "Relationship not shown",
                     "~",
                     "Averages not illustrated",
                     "~",
                     "Values not clearly shown",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unclear explanation",
                     "~",""),
                 `Full credit` = c("Reasonable values",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear illustration",
                                   "~",
                                   "Clear presentation",
                                   "~",
                                   "Clear explanation",
                                   "~",
                                   "Clear explanation",
                                   "~",
                                   ""),
                 `Possible points` = c(4, "",
                                       3,"",
                                       3,"",
                                       3,"",
                                       3,"",
                                       3,"",
                                       3,"", 
                                       3,"",
                                       25))

kable(rubric, format = "html")
```
        