---
title: "Lecture 10"
subtitle: "Multivariate Regression"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prior to this lecture, you should have read chapter 10 of
[Regression and Other Stories](https://users.aalto.fi/~ave/ROS.pdf){target="_blank"}.

```{r, include=FALSE, eval=FALSE}
library(tidyverse)
library(here) 
 
exam_data <- tibble(pretest = ceiling(rnorm(n = 500, mean = 70, sd = 7)),
                    time_studying = ceiling(rnorm(n = 500, mean = 45, sd = 15)),
                    instructor = c(rep("Jacobs", 154), 
                                   rep("Moses", 172), 
                                   rep("Krumholz", 174))) %>%
  mutate(final_score = ceiling(
           0.8*jitter(pretest) + 
           0.4*jitter(time_studying) + 
           10*(instructor == "Krumholz") + 
           3*(instructor == "Jacobs") +
           rnorm(n = 500, mean = 0, sd = 4))) %>%
  mutate(final_score = ifelse(final_score > 100, 100, final_score)) %>%
  mutate(time_studying = ifelse(time_studying < 0, 0, time_studying))

here("Examples",
     "test-scores.csv") %>%
write_csv(x = exam_data, file = .)
```

```{r, include=FALSE}
library(tidyverse)
library(here) 

exam_data <- here("Examples",
                  "test-scores.csv") %>%
  read_csv()
```

## Example data

Five hundred students enrolled in a course. Students were randomly assigned to one of three
instructors. Students took a pretest at the beginning of the term to test their initial 
knowledge of the topic. Prior to taking the final exam, students were asked how much time
they spent studying for the course each week.

Here is a scatter plot showing the relationship to students' pretest scores and their 
scores on the final exam:

```{r, echo=FALSE}
ggplot(exam_data) +
  geom_point(aes(x = pretest,
                 y = final_score)) +
  scale_x_continuous(name = "Pre-test score") +
  scale_y_continuous(name = "Final exam score") +
  theme_minimal()
```

Here is a scatter plot showing the relationship between the amount of time students
spent studying and their scores on the final exam:

```{r, echo=FALSE}
ggplot(exam_data) +
  geom_point(aes(x = time_studying,
                 y = final_score)) +
  scale_x_continuous(name = "Self-reported time spent studying per week") +
  scale_y_continuous(name = "Final exam score") +
  theme_minimal()
```

Here are the student scores for each instructor, with average indicated by 
horizonal lines.

```{r, echo=FALSE}
summary_by_instructor <- exam_data %>%
  group_by(instructor) %>%
  summarise(avg_final = mean(final_score))

ggplot(data = exam_data, aes(x = 0, y=final_score)) +
  geom_jitter() +
  geom_hline(data = summary_by_instructor, 
             aes(yintercept = avg_final))+
  facet_grid(~instructor, switch = "both") +
  scale_x_continuous(name = "") +
  scale_y_continuous(name = "Final exam score") +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.text.x = element_blank())
```

Higher test scores seem to be associated with more time spent studying, higher pretest 
scores, and with being in Krumholz's class. 

## Estimating a multiple regression model

Let's predict final exam scores based on pretest scores, weekly time spent
studying, and instructor.

```{r, echo=FALSE, out.width='100%'}
here::here("unit-5",
     "excel-multi-regression.gif") %>%
  knitr::include_graphics()
```

```{r}
model <- lm(final_score ~ pretest + time_studying + instructor,
            data = exam_data)
```

You'll get the same results in R as in Excel:

```{r}
summary(model)
```

```{r, echo=FALSE, out.width='100%'}
here::here("unit-5",
     "excel-regression.png") %>%
  knitr::include_graphics()
```

Each coefficient represents the change in the outcome (test score) that 
would be predicted based on a one-unit change in the associated predictor,
**if all other variables in the model were held constant**.