---
title: "Assignment 3"
subtitle: "Testing for relationships"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

Select one of the categories of your categorical variable
as the "reference category" (this will likely be the most
common category in your sample). 

Calculate a 95-percent confidence interval for the average value
of your outcome variable within each category of your (or each) 
categorical variable.

Calculate the difference in means between the value of the 
outcome variable in your reference category and the value
of the outcome variable in each of the other categories.

Calculate the correlation between each of your continuous
variables and the outcome variable.

## Deliverable

Submit a memo (as a PDF or html document) presenting and explaining your 
results. Include figures to illustrate your analysis as appropriate.

## Example

This example would receive full credit.

[Crashes example](https://c-voulgaris.github.io/SES-5215-F22/Examples/a3-commuting.pdf){target="_blank"}

## Rubric

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)

rubric <- tibble(Item = c("Average within each category",
                          "~",
                          "Explanation for category averages",
                          "~",
                          "Difference in means confidence interval",
                          "~",
                          "Explanation for difference in means",
                          "~",
                          "Correlation confidence interval",
                          "~",
                          "Explanation of correlation",
                          "~",
                          "Inclusion of appropriate figures",
                          "~",
                          "Total"),
                 `Half credit` = 
                   c("Unreasonable values",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unreasonable values",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unreasonable values",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Figures are not relevant",
                     "~",""),
                 `Full credit` = c("Reasonable values",
                     "~",
                     "Clear explanation",
                     "~",
                     "Reasonable values",
                     "~",
                     "Clear explanation",
                     "~",
                     "Reasonable values",
                     "~",
                     "Clear explanation",
                     "~",
                     "Figures are relevant",
                     "~",""),
                 `Possible points` = c(3, "~",
                                       4,"~",
                                       3,"~",
                                       4,"~",
                                       3,"~", 
                                       4,"~",
                                       4,"~",
                                       25))

kable(rubric, format = "html")
```
        