---
title: "Assignment 4"
subtitle: "Regression with one predictor"
output: tint::tintHtml
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions

For each predictor variable, estimate a linear regression model
predicting the outcome variable. There should be one model for
each predictor variable, and each model should include only one
predictor variable.

## Deliverable

Submit a memo (as a PDF or html document) presenting and explaining your 
results. 

## Example

This example would receive full credit.

[Commuting example](https://c-voulgaris.github.io/SES-5215-F22/Examples/a4-commuting.pdf){target="_blank"}

## Rubric

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)

rubric <- tibble(Item = c("Regression with first continuous variable",
                          "~",
                          "Interpretation of first regression coefficient",
                          "~",
                          "Interpretation of first regression model fit",
                          "~",
                          "Regression with second continuous variable",
                          "~",
                          "Interpretation of second regression coefficient",
                          "~",
                          "Interpretation of second regression model fit",
                          "~",
                          "Regression with categorical variable",
                          "~",
                          "Interpretation of categorical coefficients",
                          "~",
                          "Interpretation of categorical model fit",
                          "~",
                          "Total"),
                 `Half credit` = 
                   c("Unreasonable results",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unreasonable results",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unreasonable results",
                     "~",
                     "Unclear explanation",
                     "~",
                     "Unclear explanation",
                     "~",""),
                 `Full credit` = c("Reasonable results",
                     "~",
                     "Clear explanation",
                     "~",
                     "Clear explanation",
                     "~",
                     "Reasonable results",
                     "~",
                     "Clear explanation",
                     "~",
                     "Clear explanation",
                     "~",
                     "Reasonable results",
                     "~",
                     "Clear explanation",
                     "~",
                     "Clear explanation",
                     "~",""),
                 `Possible points` = c(2, "~",
                                       3,"~",
                                       3,"~",
                                       2,"~",
                                       3,"~", 
                                       3,"~",
                                       3,"~",
                                       3,"~",
                                       3,"~",
                                       25))

kable(rubric, format = "html")
```
        